# ggplot2 (60 minutos)

Alguns links

[The R Graph Gallery](https://r-graph-gallery.com/)

[120 registered extensions available to explore](https://exts.ggplot2.tidyverse.org/gallery/)

[link 1: patchwork](https://patchwork.data-imaginist.com/articles/guides/assembly.html)

[link 2: patchwork](https://patchwork.data-imaginist.com/articles/patchwork.html)


```{r}
ls("package:ggplot2")
```

## Primeiros passos

```{r}
ggplot(dados)

ggplot(dados, aes(x = cty, y = hwy))

# Alternativas
ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point()

ggplot(dados) + 
  geom_point(aes(x = cty, y = hwy))

ggplot() + 
  geom_point(data = dados, aes(x = cty, y = hwy))
# Fim 

ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(colour = "red")

ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(colour = "red", size = 6)

ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(colour = "red", size = 6, shape = 10)

ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(colour = "red", size = 6, shape = 10)+
  labs(x = "cty (city miles per gallon hwy)", 
       y = "hwy (highway miles per gallon)", 
       title = "Pensar em algum título...", 
       subtitle = "Escrever alguma coisa")
```

### Mais detalhes sobre geom_point {-}

> geom_point() understands the following aesthetics (required aesthetics are in bold):

  - x

  - y

  - alpha

  - colour

  - fill

  - group

  - shape

  - size

  - stroke
  
## colour, shape, fill, alpha

```{r}
ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point()

ggplot(dados, aes(x = cty, y = hwy, col = factor(year))) + 
  geom_point() + 
  labs(col = "year")

ggplot(dados, aes(x = cty, y = hwy, alpha = factor(year))) + 
  geom_point() + 
  labs(alpha = "year")

ggplot(dados, aes(x = cty, y = hwy, size = factor(year))) + 
  geom_point() + 
  labs(size = "year")

# Alternativa
ggplot(dados, aes(x = cty, y = hwy, col = cty < 20)) + 
  geom_point() + 
  labs(col = "year")

# Erro comum
ggplot(dados, aes(x = cty, y = hwy, col = "red")) + 
  geom_point()+
  labs(col = "year")

ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(col = "red")+
  labs(col = "year")
# Fim  Erro comum

ggplot(dados, aes(x = manufacturer, y = hwy, fill = factor(year))) + 
  geom_col(position = "dodge") + 
  labs(fill = "year") +
  theme(axis.text.x = element_text(angle = 45))

ggplot(dados, aes(x = cty, y = hwy, shape = factor(year))) + 
  geom_point(col = "red") + 
  labs(shape = "year")


ggplot(dados, aes(x = cty, y = hwy, size = factor(class))) + 
  geom_point() + 
  labs(size = "year")

ggplot(dados, aes(x = cty, y = hwy, size = factor(class), col = factor(class))) + 
  geom_point() + 
  labs(size = "year", 
       col = "class")

```

## point, smooth, boxplot, histogram


```{r}
glimpse(dados)

v1 <- ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(col = "blue")
v1

v2 <- ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_point(col = "blue")+
  geom_smooth(method = mgcv::gam,
              formula = y ~ s(x, bs = "cs") ,
              col = "red", 
              se = FALSE)
v2

v3 <- ggplot(dados, aes(x = cty)) + 
  geom_boxplot(fill = "red")
v3

v4 <- ggplot(dados, aes(x = cty)) + 
  geom_histogram(bins = 10, fill = "red", col = "blue", lwd=2)
v4

# Alternativas
ggplot(dados, aes(x = cty)) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = 10, fill = "yellow", col = "red") +
  geom_density(col = "blue", lwd =3)
```

## gridExtra e patchwork


Alguns links

[link 1: patchwork](https://patchwork.data-imaginist.com/articles/guides/assembly.html)

[link 2: patchwork](https://patchwork.data-imaginist.com/articles/patchwork.html)

```{r}
# gridExtra
grid.arrange(v1, v2, v3, v4) 

# patchwork
v1 + v2
v1 | v2
v1 / v2


v1 + v2 + v3
v1 + (v2 + v3)
v1 | (v2 / v3)
v1 / (v2 + v3)

v1 + v2 + v3 + v4 
v1/(v2+v3+v4)
v1  + (v2 + v3 + v4)
v1  + v2 + (v3 + v4)
(v1 | v2 | v3) / v4
```


## bar, col, density, density2d

```{r}
v5 <- ggplot(dados , aes(x = manufacturer)) + 
  geom_bar()+ 
  theme(axis.text.x = element_text(angle = 45))
v5

v6 <- ggplot(dados , aes(x = manufacturer, y = cty)) + 
  geom_col()+
  theme(axis.text.x = element_text(angle = 45))
v6

v7 <- ggplot(dados , aes(x = cty)) + 
  geom_density()
v7

v8 <- ggplot(dados, aes(x = cty, y = hwy)) + 
  geom_density2d()
v8

(v5+v6)/ (v7 + v8)
```

## facet_grid, facet_grid

```{r}
p1<- ggplot(dados, aes(x = cty, y = hwy)) +
  geom_point()
p1

p1 + facet_grid(rows = vars(cyl))
p1 + facet_grid(cols = vars(cyl))
p1 + facet_grid(~cyl)
p1 + facet_grid(rows = vars(year), cols =vars(cyl))
p1 + facet_grid(year~cyl)

p1 + facet_wrap(year~cyl)
p1 + facet_wrap(year~cyl, ncol = 4)
```


## stat_function

```{r}
a<- -3 # média
b<- 4  # desv. padrão
ggplot(data.frame(x = c(a - 3*b, a + 3*b)), aes(x)) + 
  stat_function(fun = dnorm, args = list(mean = a, sd = b))+
  geom_vline(xintercept = c(a - 3*b, a, a + 3*b), col = "red", lty = 2)+
  theme_minimal()
```

## theme_

```{r}
a1<- p1 + theme_bw() + labs(title = "theme_bw()")
a2<- p1 + theme_classic() + labs(title = "theme_classic()")
a3<- p1 + theme_light() + labs(title = "theme_light()")
a4<- p1 + theme_minimal() + labs(title = "theme_minimal()")

a1 + a2 + a3 + a4
```

## plotly


[plotly cran](https://cran.r-project.org/web/packages/plotly/index.html)

[Interactive web-based data visualization with R, plotly, and shiny](https://plotly-r.com/)

[Plotly R Open Source Graphing Library](https://plotly.com/r/)

```{r, eval=FALSE}
ggplotly(v1)
ggplotly(v2)
ggplotly(v4)
ggplotly(v5)
```


## esquisse

[esquisse](https://cran.r-project.org/web/packages/esquisse/vignettes/get-started.html)

[esquisse + shiny](https://cran.r-project.org/web/packages/esquisse/vignettes/shiny-usage.html)
